---
title: 'Homework #3'
author: "Vincent Tam"
date: "October 9, 2018"
output:
  github_document: default
---
```{r Library Package Setup and Global Options, echo = FALSE}
## Prepare Setup to Answer Problem 3
## Install Libraries
library(tidyverse)
library(ggplot2)
library(forcats)
library(knitr)
library(ggridges)
library(janitor)
## install.packages("devtools") Already Installed but do not want to do it constantly
## Upload Datasets
library(p8105.datasets)
data(brfss_smart2010)
```
### Problem 1 - Behavioral Risk Factors Surveillance System (BRFSS) Data
```{r Clean and Format Data to Prepare for Answering Questions, echo = FALSE}
## Clean, Filter, Select, Mutate, and Separate Variables
brfss_data = 
  brfss_smart2010 %>%
## format the data to use appropriate variable names
  janitor::clean_names() %>%
## focus on the “Overall Health” topic
  filter(topic == "Overall Health") %>%
## include only responses from “Excellent” to “Poor”
  filter(response == "Excellent" | response == "Very good" | response == "Good" | response == "Fair" | response == "Poor") %>%
## organize responses as a factor taking levels ordered from “Excellent” to “Poor”
  mutate(cleaned_response = as.factor(response)) %>%
  ## organize and select relevant columns for questions
  select(year, locationdesc, topic, response, data_value) %>%
  separate(locationdesc, into = c("state", "county"), sep = " - ")
```
Answers to Homework Questions
```{r Answering Questions, echo = FALSE}
brfss_data_2002 =
  filter(brfss_data, year == "2002") %>%
  distinct(state, county) %>%
  count(state) %>%
  filter(n == 7)
## In 2002, which states were observed at 7 locations?
brfss_data_2002
```
In 2002, Connecticut, Floriday, and North Carolina were states observed at 7 locations
```{r Spaghetti Plotting, echo = FALSE}
## Spaghetti Plots
brfss_plot = 
  brfss_data %>%
  distinct(year, state, county) %>%
  select(year, state, county) %>%
  group_by(year, state) %>%
  mutate(count = n())
ggplot(brfss_plot, aes(x = year, y = count, color = state)) + geom_point() + geom_line()
```
"The Spaghetti Plot shows the number of locations in each state from 2002 to 2010. The vast majority of states are observed to have 10 locations or less. A few exceeded 10 but had less than 20. A peak in 2007 shows 40+ locations and a similar 40+ peak was observed in 2010."
```{r Making Table of Excellent Responses in NY State, echo = FALSE}
brfss_table =   
  brfss_data %>%
  filter(year == '2002' | year == '2006' | year == '2010') %>%
  filter(state == 'NY') %>%
  filter(response == 'Excellent') %>%
  mutate(data_value_numeric = as.numeric(data_value)) %>%
  group_by(year) %>%
  summarize(brfss_NYexcellentsd = sd(data_value_numeric), brfss_NYexcellentmean = mean(data_value_numeric))
brfss_table
```
The standard deviations (SD) and means of years 2002, 2006, and 2010, were calculated for the proportions of 'Excellent' responses in NY state. 2002 showed an SD of 4.49 and a mean of 24.04. 2006 had an SD of 4.00 and a mean of 22.53. Year 2010 had an SD of 3.57 and a mean of 22.70. The means and SDs of the three years appear to be similar and consistent.
```{r 5-Panel Plot Comparing Proportions of Responses, warning = FALSE}
brfss_proportions = 
  brfss_data %>%
  group_by(year, state, response) %>%
  summarize(brfss_excellentmean = mean(as.numeric(data_value)), 
            brfss_verygoodmean = mean(as.numeric(data_value)), 
            brfss_goodmean = mean(as.numeric(data_value)),
            brfss_fairmean = mean(as.numeric(data_value)),
            brfss_poormean = mean(as.numeric(data_value))) %>%
  gather(key = cleaned_response, value = proportion, brfss_excellentmean:brfss_poormean) %>% 
  mutate(response = forcats::fct_relevel(response, c("Poor", "Fair", "Good", "Very good", "Excellent"))) 
ggplot(brfss_proportions, aes(x = year, y = proportion, fill = response)) + geom_violin() + facet_grid(~response) + theme(axis.text.x = element_text(angle = 90)) + labs(title = "Average Proportion per Response Category"  )
```
The side-by-side plots of the response categories: Excellent, Very Good, Good, Fair, and Poor, show relatively consistent distribution in their respective ranges through the years examined. The response Very Good represented the highest proportion of responses every year.

## Problem 2 - Instacart Data
```{r Problem 2 Instacart Data, echo = FALSE}
## Uploading and Cleaning Data
data(instacart)
instacart_cleaned = 
  instacart %>%
  janitor::clean_names() 
## Number of Aisles in Dataset
  length(unique(instacart_cleaned$aisle)) 
## Top 5 Most Ordered Items
most_ordered = instacart_cleaned %>%
  group_by(aisle) %>%
  summarize(n = n()) %>%
  filter(min_rank(desc(n)) < 6)
most_ordered
```
There are 134 different aisles. The top five aisles where the most food are ordered from, are:  
1. Fresh Vegetables  2. Fresh Fruites  3. Packaged Vegetables Fruits  4. Yogurt  5. Packaged Cheese  
```{r Plot of number of items ordered in each aisle, ordered and organized sensibly, echo = FALSE}
## Organizing Data from Least to Most
instacart_cleaned %>% 
  group_by(aisle_id) %>%
  summarize(n = n()) %>%
  mutate(aisle_id = forcats::fct_reorder(as.factor(aisle_id), n)) %>% 
## Plotting Number of Items in Each Aisle from Least to Most
ggplot(aes(x = aisle_id, y = n)) + geom_bar(stat = "identity") + labs(title = "Number of Items per Aisle",    x = "Aisle ID", y = "Number of Items") + theme(axis.text.x = element_text(angle = 45, size = 5))
```
The plot above shows approximately 150,000+ items stocked in the aisles in the dataset. Fresh Vegetables, Aisle 83 hold the most items while Beauty, Aisle 132, contains the least. 
```{r Most Popular Food Items, echo = FALSE}
## Cleaning Data for Specific Aisles
popular_item = 
  instacart %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  select(aisle, product_name) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
## Sorting Most Popular Items
popular_item_sort = 
  popular_item %>%
  group_by(aisle) %>% 
  top_n(1) %>% 
  knitr::kable(digits = 1)
popular_item_sort
```
The table above shows the most popular items for their respective aisles. In Packaged Vegetables Fruits, Organic Baby Spinach had 9784 orders, 20x more than the most popular Baking Ingredient, Light Brown Sugar (499 orders). Meanwhile, Dog Food Care's most popular item is Snack Sticks Chicken & Rice Recipe Dog Treats, with 30 orders. 
```{r Pink Lady Apples & Coffee Ice Cream, echo = FALSE}
## Cleaning Data
apples_icecream = 
  instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  spread(key = product_name, value = mean_hour) %>% 
  mutate(order_dow = recode(order_dow, `0` = "Sunday", `1` = "Monday", `2` = "Tuesday", `3` = "Wednesday", `4` = "Thursday", `5` = "Friday", `6` = "Saturday")) %>%
   knitr::kable(digits = 1)
## Chart
apples_icecream
```
The table above shows the average time of day where people order pink-lady apples and coffee ice cream is generally during mid-day. However, coffee ice cream is ordered the latest on Tues, Wed, and Thurs. Pink-lady apples  This habit appears to tend to be ordered the latest on Wednesdays.
## Problem 3 - New York NOAA Data
```{r}
data(ny_noaa)
```
```{r}
nynoaa_cleaned = 
  ny_noaa %>% 
  janitor::clean_names() %>%
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(tmin = as.numeric(tmin)/10, tmax = as.numeric(tmax)/10)
```
```{r}
nynoaa_snow = 
  nynoaa_cleaned %>%
  group_by(snow) %>%
  summarize(n = n()) %>%
  top_n(3)
nynoaa_snow
```
```{r}
maxtempJanJul_plot =
  nynoaa_cleaned %>%
  na.omit() %>%
  select(year, month, id, tmin, tmax) %>%
  filter(month == "01" | month == "07") %>%
    group_by(year, month, id) %>%
  summarize(tmax_mean = mean(tmax)) %>%
  ggplot(aes(x = year, y = tmax_mean)) + geom_boxplot() + facet_grid(.~month) + theme(axis.text.x = element_text(size = 9, angle = 90)) + labs(title = "Comparison of Average Max Temps in Celcius between January and July from 1981 to 2010", x = "Years", y = "Temperature (Celcius)" )
maxtempJanJul_plot
```
```{r}
plot1_minmax =
  nynoaa_cleaned %>%
  na.omit() %>%
  select(tmin, tmax) %>%
  ggplot(aes(x = tmin, y = tmax )) + geom_hex() + labs(title = "Minimum Temperatures vs Maximum Temperatures", x = "Minimum Temp", y = "Maximum Temp") + theme(axis.text.x = element_text(size = 9, angle = 90))
plot1_minmax
```
```{r}
plot2_snowfall =
  nynoaa_cleaned %>%
  na.omit() %>%
  filter(snow > 0 & snow < 100) %>%
  ggplot(aes(x = year, y = snow )) + geom_boxplot() + labs(title = "Distribution of Snowfall Values by Years", x = "Year", y = "Snowfall Valuesp") + theme(axis.text.x = element_text(size = 9, angle = 90))
plot2_snowfall
```